---
title: "R Notebook"
output: html_notebook
---

*1. Data Cleaning*
- reads police_incidents.csv and loads only required attributes into a dataframe object
```{r}
library(tibble)
library(dplyr)
library(readr)
dallas_incidents<-read_csv('Documents/dallas/Police_Incidents.csv')
#as_tibble(dallas_incidents)
dallas=select(dallas_incidents,1,3,5,6,7,9,38,43,44,46,47,48,50,53,54,69,75,76,78,95,96,97)
```
*2. Data Pre-processing*
- transforms the attribute 'call received date time' string to R datetime object and sort them in ascending order  
```{r}
library(lubridate)
library(magrittr)
dallas%>%
  mutate(`Call Received Date Time`=as_datetime(mdy_hms(`Call Received Date Time`)))%>%
  arrange(`Call Received Date Time`)
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
