---
title: "Ashish's R Notebook"
output: html_notebook
---


#Importing the Dataset

```{r}

library(tidyverse)

PIdata <- read_csv("Crimedataset/Police_Incidents.csv")

```
#Examining the dataset
```{r}

class(PIdata)

```
```{r}

glimpse(PIdata)

```
#1.Data Cleaning and Preprocessing:

#Selecting columns releveant for Analysis and Modifying some of them
```{r}

as_tibble(PIdata)#Converting the Dataset from a dataframe into a tibble

Dallas = select (PIdata, 1, 3, 5, 6, 7, 9, 13, 14, 15, 38, 43, 44, 46, 47, 48, 50, 53, 54, 69, 75, 76, 78, 95, 96, 97)

Dallas

```
# Took Columns 13, 14 ,15 too because:
#13- gives an idea about in which area, the crime was reported
#14- gives an idea about police station/beat was investigating the crime/wrote down the incident report
#15- can divulge information regarding which part of the city, most incidents occur. for example- Northeast, east , west etc. 
#Note: I want to try and create Some visaulizations using these columns.

```{r}

library(lubridate)

library(magrittr)

Select(Dallas, Hate Crime)










